### 2023-05-08 paper.Rmd
WIP

---
title: "Cars Auctions"
author: "Erik Nuroyan, Narek Sahakyan"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

\newpage

# Introduction

Intro text

# Eda

eda text